Higher-Order unification for free!

### 2024-04-18 paper/main.tex
reformulate lemma 7.7 + proof

\noindent In order to accommodate for the implementation of heuristics we
shall relax these failure conditions in~\cref{sec:heuristics}.

@@

\begin{proof}[Proof sketch]
  Let $l$ a \linkbeta in the store \linkStore. If $l$ is activated by
  \progBetaRH{}, then it disappears from \linkStore and \elpiIn{progress}
  terminates. Otherwise, the \rhs of $l$ is made by a variable applied to $m$
  arguments. At each activation of \progBetaLL, $l$ is replaced by a new
  \linkbeta $l^1$ having $m-1$ arguments. At the $m^{th}$ iteration, the
  \linkbeta $l^m$ has no more arguments and is removed from \linkStore.
  Note that at the $m^{th}$ iteration, $m$ new \linketa have been added to
  \linkStore, however, by \cref{lemma:prog-eta-terminates}, the algorithm
  terminates. Finally \progBetaFail guarantees termination since 
  %it makes 
  %\elpiIn{progress}
  it immediately fails.
%
  % \Cref{def:progBetaRH} makes \elpiIn{progress} terminates, since it makes a
  % \linkbeta disappear from \linkStore. On the other hand, \cref{def:progBetaLL},
  % creates each time a new \linketa and replace the old \linkbeta with a new one.
  % This progression can however triggered at most $m$ times, since each new
  % \linkbeta as a smaller list of applied variables. At the $m^{th}$ progression,
  % the \linkbeta is removed from \linkStore which is now filled by $m$ new
  % \linketa. By \cref{lemma:prog-eta-terminates}, we know that \elpiIn{progress}
  % terminates if \linkStore is made by only \linketa and therefore
  % \elpiIn{progress} still terminates.
  %\noteE[inline]{is it ok?}
  % \noteE[inline]{funziona. per essere più precisi io parlerei di ordine lessicografico (tipico ordine ben fondato usato per dimostrare terminazione). Nl nostro caso è la tripla (argomenti extra dei beta, numero di beta, numero di eta).}
\end{proof}

@@

\begin{theorem}[Fidelity in \Fo]
  The introduction of \linkbeta guarantees \fullref{prop:fidelity-recovery}
  if \Uo{} fails outside \llambda.
\end{theorem}

\begin{proof}[Proof sketch]
  Let $\foUnifPb_i$ be the first problem in \foUnifPb not in \llambda and
  le $\sigma$ be the substitution obtained solving $\hoUnifPb_1\ldots\hoUnifPb_{i-1}$.
  If $\sigma\hoUnifPb_i$ is in \llambda, then each variable that the
  compiler used to replace a problemat term is assigned and hence
  by \cref{def:progBetaLL,def:progBetaRH} the associated \linkbeta
  been solved and removed. Note that in this case all
  calls to \Ue{} are between terms in \llambda and by \cref{lemma:fidelity-maybeeta}
  fidelity is guaranteed. If $\sigma\hoUnifPb$ is still in \notllambda,
  then by \cref{def:progBetaFail} unification fails as the
  corresponding unification in \Fo{} would.
\end{proof}

>>>>

We will relax this last failure condition to accommodate for the
implementation of heuristics in~\cref{sec:heuristics}.

Finally, in order to correctly correlate a variable appearing in the \lhs of a
\linketa and in the \rhs of a \linkbeta, we should introduce the following
\linketa progression rule.

\newcommand{\progressetaright}{\emph{$\eta$-progress-\rhs}\xspace}
\begin{definition}[\progressetaright]\label{def:progressetaright} A link
\linketaM{\Gamma}{X}{T} is removed from \linkStore when either 1)
$\testmaybeeta~T$ does not hold (anymore) or 2) by $\eta$-contracting $T$ to
$T'$ where $T'$ does not start with the \elpiIn{lam} constructor. In the
first case, $X$ is unified with $T$, and in the second one, $X$ is unified with
$T'$ (under the context $\Gamma$).
\end{definition}

The idea behind this is to instantiate the variable of a \linketa whenever its
\rhs is for sure a \wellb term.

@@

\begin{proof}[Proof sketch]
  Let $l$ a \linkbeta in the store \linkStore. If $l$ is activated by
  \progBetaRH{}, then it disappears from \linkStore and \elpiIn{progress}
  terminates. Otherwise, the \rhs of $l$ is made by a variable applied to $m$
  arguments. At each activation of \progBetaLL, $l$ is replaced by a new
  \linkbeta $l^1$ having $m-1$ arguments. At the $m^{th}$ iteration, the
  \linkbeta $l^m$ has no more arguments and is removed from \linkStore.
  Note that at the $m^{th}$ iteration, $m$ new \linketa have been added to
  \linkStore, however, by \cref{lemma:prog-eta-terminates}, the algorithm
  terminates. \progBetaFail guarantees termination since 
  %it makes 
  %\elpiIn{progress}
  it immediately fails and \progressetaright terminates since it executes 
  terminating instructions and if its permises succeed, ther considered
  \linketa is removed from \linkStore.
%
  % \Cref{def:progBetaRH} makes \elpiIn{progress} terminates, since it makes a
  % \linkbeta disappear from \linkStore. On the other hand, \cref{def:progBetaLL},
  % creates each time a new \linketa and replace the old \linkbeta with a new one.
  % This progression can however triggered at most $m$ times, since each new
  % \linkbeta as a smaller list of applied variables. At the $m^{th}$ progression,
  % the \linkbeta is removed from \linkStore which is now filled by $m$ new
  % \linketa. By \cref{lemma:prog-eta-terminates}, we know that \elpiIn{progress}
  % terminates if \linkStore is made by only \linketa and therefore
  % \elpiIn{progress} still terminates.
  %\noteE[inline]{is it ok?}
  % \noteE[inline]{funziona. per essere più precisi io parlerei di ordine lessicografico (tipico ordine ben fondato usato per dimostrare terminazione). Nl nostro caso è la tripla (argomenti extra dei beta, numero di beta, numero di eta).}
\end{proof}

@@

\begin{theorem}[Fidelity in \Fo]
  The introduction of \linkbeta guarantees \fullref{prop:fidelity-recovery}
  if \Uo{} falls outside \llambda.
\end{theorem}

\begin{proof}[Proof sketch]
  Let $\foUnifPb_i$ be the first problem in \foUnifPb, such that it is not in \llambda and le
  $\sigma$ be the substitution obtained solving
  $\hoUnifPb_1\ldots\hoUnifPb_{i-1}$. After the execution of $\hstep_{i-1}$,
  each variable $X$ corresponding to a \maybeeta subterm in the original \Fo 
  unification problem, is instantiated if it is known for sure
  that $X$ is no more a \maybeeta. If $\sigma\hoUnifPb_i$ is in \llambda, then 
  % each variable that the
  % compiler used to replace a ``problematic'' term is assigned and hence
  by \cref{def:progBetaLL,def:progBetaRH} the associated \linkbeta
  is solved and removed. Note that, in this case, all
  calls to \Ue{} are between terms in \llambda and by \cref{th:fidelity-beta}
  fidelity is guaranteed. If $\sigma\hoUnifPb$ is still in \notllambda,
  then by \cref{def:progBetaFail} unification fails as the
  corresponding unification in \Fo{} would.
\end{proof}
